Consistent, and weak, ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-08-22 manuscript/usses_water_ms.Rmd
more abstract edits

---
title: Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe
preprint: false
author: 
  - name: Andrew T. Tredennick
    affiliation: 1
    corresponding: true
    email: atredenn@gmail.com
  - name: Andrew R. Kleinhesselink
    affiliation: 2
  - name: Bret Taylor
    affiliation: 3
  - name: Peter B. Adler
    affiliation: 1
affiliation:
  - code: 1
    address: Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322
  - code: 2
    address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California 90095
  - code: 3
    address: United States Department of Agriculture, Agriculture Research Station, U.S. Sheep Experiment Station, Dubois, Idaho 83423
abstract: |
  **Background.** Precipitation is predicted to become more variable in the western United States, meaning years of above and below average precipitation will become more common. Periods of extreme precipitation are major drivers of ecosystem functioning in water limited grasslands, but how ecosystems respond to precipitation may change as the duration of above and below average periods increases. Changes in ecosystem functional response could reflect compensatory changes in species composition or species reaching physiological thresholds at extreme precipitation levels.

@@

  **Discussion.** Despite some evidence that ecosystem functional response saturated under extreme wet conditions, the response of ANPP to soil moisture and community compositional changes in this sagebrush steppe system were consistently small. The similarity, or difference, of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, but instead reflects the weak responses of individual species. Such responses may be due to bet-hedging strategies where species respond weakly to even extreme levels of soil moisture to ensure low variance of long-term success.
header-includes: >
  \usepackage{setspace}
  \usepackage{todonotes}
  \usepackage{rotating}
  \usepackage{color, soul}
  \usepackage{sectsty}
  \usepackage{bm,mathrsfs}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
output: 
  rticles::peerj_article:
    number_sections: false
    base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
---

@@

The results of our five year experiment in a sagebrush steppe conform to two of our four predictions, depending on treatment.
Under chronic drought, neither ecosystem functional response nor community composition changed (Fig. 1, top left).
But under irrigation, ecosystem functional response was different from the control treatment while community composition remained unchanged (Fig. 1, top right).
Thus, we found some evidence for a saturating response at high water availability (Fig. 3), but the response of ANPP to soil moisture was weak overall, regardless of treatment (Table 1).
<!---
The results of our five year experiment in a sagebrush steppe conform to the first of our four predictions: neither ecosystem functional response nor community composition changed with chronic alteration in water availability (Fig. 1, top left).
--->

>>>>

---
title: Consistent, and weak, ecosystem functional response across precipitation extremes in a sagebrush steppe
preprint: false
author: 
  - name: Andrew T. Tredennick
    affiliation: 1
    corresponding: true
    email: atredenn@gmail.com
  - name: Andrew R. Kleinhesselink
    affiliation: 2
  - name: Bret Taylor
    affiliation: 3
  - name: Peter B. Adler
    affiliation: 1
affiliation:
  - code: 1
    address: Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322
  - code: 2
    address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California 90095
  - code: 3
    address: United States Department of Agriculture, Agriculture Research Station, U.S. Sheep Experiment Station, Dubois, Idaho 83423
abstract: |
  **Background.** Precipitation is predicted to become more variable in the western United States, meaning years of above and below average precipitation will become more common. Periods of extreme precipitation are major drivers of ecosystem functioning in water limited grasslands, but how ecosystems respond to precipitation may change as the duration of above and below average periods increases. Changes in ecosystem functional response could reflect compensatory changes in species composition or species reaching physiological thresholds at extreme precipitation levels.

@@

  **Discussion.** Despite some evidence that ecosystem functional response saturated under extreme wet conditions, the response of ANPP to soil moisture was consistintly weak and community composition was remarkably stable. The similarity, or difference, of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, but instead reflects the weak responses of individual species. Such responses may be due to bet-hedging strategies where species respond weakly to even extreme levels of soil moisture to ensure low variance of long-term success.
header-includes: >
  \usepackage{setspace}
  \usepackage{todonotes}
  \usepackage{rotating}
  \usepackage{color, soul}
  \usepackage{sectsty}
  \usepackage{bm,mathrsfs}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
output: 
  rticles::peerj_article:
    number_sections: false
    base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
---

@@

The results of our five year experiment in a sagebrush steppe conform to two of our four predictions, depending on treatment.
Under chronic drought, neither ecosystem functional response nor community composition changed (Fig. 1, top left).
But under irrigation, ecosystem functional response was different from the control treatment while community composition remained unchanged (Fig. 1, top right).
Thus, we found some evidence for a saturating response at high water availability (Fig. 3A, right panel).
However, the response of ANPP to soil moisture was consitently weak, regardless of treatment (Table 1).
<!---
The results of our five year experiment in a sagebrush steppe conform to the first of our four predictions: neither ecosystem functional response nor community composition changed with chronic alteration in water availability (Fig. 1, top left).
--->
